### 2022-03-29 chapters/drto/article.tex
modifs post-defense part III
modifs part IV

We studied the transmission patterns of SCTLD between the Marquesas and the DRTO by using the hydro-epidemiological model of \cite{dobbelaere2020coupled}. This model relies on the multi-scale ocean model SLIM\footnote{\url{ https://www.slim-ocean.be}} to simulate the hydrodynamics over the entirety of FCR. The computational domain includes the Florida Straits as well as part of the GoM (Fig. \ref{fig:fig1_drto}a). SLIM is coupled with the large-scale ocean model HYCOM in order to correctly represent the dynamics of the FC which is a major western boundary current that flows out of the GoM, as a continuation of the Loop Current, before joining the Gulf Stream. SLIM uses an unstructured mesh whose resolution can be locally increased in order to accurately represent fine-scale flow features. Here we used the same mesh as \cite{dobbelaere2020coupled}, composed of $7\times 10^5$ elements, and with resolution that reaches 100 m over reefs and along the coastline (Fig \ref{fig:fig1_drto}a). By using such a fine resolution, we can explicitly represent circulation and resolve tides at the reef scale, such as around the Marquesas island group (Fig. \ref{fig:fig1_drto}a, inset), while also representing the large-scale flow features within the FC (Fig 1b and c). Further details on the model formulation and validation are provided by \cite{frys2020fine}.

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{chapters/drto/figures/fig4.png}
    \caption{ (a) Vulnerability indicator value of all sub-reefs of the DRTO in November 2020. Eastern reefs of the DRTO appear most vulnerable to the disease. Locations where the SCTLD was observed in the DRTO are indicated by red stars (b) Source indicator value in November 2020 for every sub-reef of the Marquesas and the Western Marquesas. Northwestern reefs of the Western Marquesas and southern reefs between the Marquesas and the Western Marquesas show a higher transmission potential to the DRTO. Main islands are shown in dark grey, separation between the Marquesas and the Western Marquesas is indicated by a dashed line.}
    \label{fig:fig4_drto}
\end{figure}

\section{Discussion and conclusion}
Our results show that the eddies generated by the LC/FC system strongly influence the connectivity between the Marquesas and the DRTO. Mesoscale eddies modify the westward coastal currents by breaking them through strong cross-shore flow regime \citep{kourafalou2012florida}, as illustrated for October 2020 in Fig. \ref{fig:fig1_drto}. \cite{lee1994evolution} hypothesized that such cross-shore flow could impact the transport of larvae spawned in the Florida Keys and cause their retention on the West Florida Shelf for periods on the order of 6 months. As Tortugas eddies may remain stationary near the DRTO for up to 120 days \citep{fratantoni1998influence}, they may thus act as barriers for the exchanges between the Marquesas and the DRTO for extended periods of time. Moreover, the passage of mesoscale eddies was identified as one of the physical mechanisms likely to deliver larvae from the DRTO to the Upper Keys, therefore opposing exchanges from the Marquesas to the DRTO. Our results confirm the role of mesoscale eddies as potential barriers to connectivity between these two regions as most months during which Tortugas eddies were present correspond to a vanishing connectivity indicator. The fact that no eddies were observed at the beginning of our simulation might be related to the identification method, based solely on visual assessment of the residual currents. A comprehensive identification of all Tortugas eddies during our simulated period would require the analysis of sea surface temperature and elevation, which is beyond the scope of the present study. While our approach might miss some Tortugas gyres, it filters out submesoscale eddies with lifespan of the order of several days. Consequently, when horizontally elongated mesoscale cyclonic streamlines are present in the computed monthly-averaged currents, they can be identified as Tortugas eddies with sufficient certainty.

@@

\begin{subappendices}
	\section{Validation of the model}

>>>>

We studied the transmission patterns of SCTLD between the Marquesas and the DRTO by using the hydro-epidemiological model of \cite{dobbelaere2020coupled}. This model relies on the multi-scale ocean model SLIM\footnote{\url{ https://www.slim-ocean.be}} to simulate the hydrodynamics over the entirety of FCR. The computational domain includes the Florida Straits as well as part of the GoM (Fig. \ref{fig:fig1_drto}a). SLIM is coupled with the large-scale ocean model HYCOM in order to correctly represent the dynamics of the FC which is a major western boundary current that flows out of the GoM, as a continuation of the Loop Current, before joining the Gulf Stream. SLIM uses an unstructured mesh whose resolution can be locally increased in order to accurately represent fine-scale flow features. Here we used the same mesh as \cite{dobbelaere2020coupled}, composed of $7\times 10^5$ elements, and with resolution that reaches 100 m over reefs and along the coastline (Fig \ref{fig:fig1_drto}a). By using such a fine resolution, we can explicitly represent circulation and resolve tides at the reef scale, such as around the Marquesas island group (Fig. \ref{fig:fig1_drto}a, inset), while also representing the large-scale flow features within the FC (Fig 1b and c). Further details on the model are provided by \cite{frys2020fine}. \modif{The model results were validated against tide gauge and ADCP measurements and showed good agreement with observations (see appendix \ref{drto:validation})}.

@@

\begin{figure}
    \centering
    \includegraphics[width=.6\textwidth]{chapters/drto/figures/vulnerability_new.png}
    \includegraphics[width=.9\textwidth]{chapters/drto/figures/source_new.png}
    \caption{ (a) Vulnerability indicator value of all sub-reefs of the DRTO in November 2020. Eastern reefs of the DRTO appear most vulnerable to the disease. Locations where the SCTLD was observed in the DRTO are indicated by red stars (b) Source indicator value in November 2020 for every sub-reef of the Marquesas and the Western Marquesas. Northwestern reefs of the Western Marquesas and southern reefs between the Marquesas and the Western Marquesas show a higher transmission potential to the DRTO. Main islands are shown in dark grey, separation between the Marquesas and the Western Marquesas is indicated by a dashed line.}
    \label{fig:fig4_drto}
\end{figure}

\section{Discussion and conclusion}
Our results show that the eddies generated by the LC/FC system strongly influence the connectivity between the Marquesas and the DRTO. Mesoscale eddies modify the westward coastal currents by breaking them through strong cross-shore flow regime \citep{kourafalou2012florida}, as illustrated for October 2020 in Fig. \ref{fig:fig1_drto}. \cite{lee1994evolution} hypothesized that such cross-shore flow could impact the transport of larvae spawned in the Florida Keys and cause their retention on the West Florida Shelf for periods on the order of 6 months. As Tortugas eddies may remain stationary near the DRTO for up to 120 days \citep{fratantoni1998influence}, they may thus act as barriers for the exchanges between the Marquesas and the DRTO for extended periods of time. Moreover, the passage of mesoscale eddies was identified as one of the physical mechanisms likely to deliver larvae from the DRTO to the Upper Keys, therefore opposing exchanges from the Marquesas to the DRTO. Our results confirm the role of mesoscale eddies as potential barriers to connectivity between these two regions as most months during which Tortugas eddies were present correspond to a vanishing connectivity indicator. The fact that no eddies were observed at the beginning of our simulation might be related to the identification method, based solely on visual assessment of the residual currents. \modif{Alternative methods to identify the presence of eddies could have been the computation of the Okubo-Weiss parameter \citep{okubo1970horizontal,weiss1991dynamics} or the extraction of the FC position as the 20$^\circ$C isotherm at 150 m from HYCOM temperature fields, following the approach of \cite{kourafalou2012florida}}. A comprehensive identification of all Tortugas eddies during our simulated period would require the analysis of sea surface temperature and elevation, which is beyond the scope of the present study. . While our approach might miss some Tortugas gyres, it filters out submesoscale eddies with lifespan of the order of several days. Consequently, when horizontally elongated mesoscale cyclonic streamlines are present in the computed monthly-averaged currents, they can be identified as Tortugas eddies with sufficient certainty.

@@

\begin{subappendices}
	\section{Validation of the model}\label{drto:validation}

### 2022-03-29 chapters/intro/intro.tex
modifs part IV

\section{Modeling hydrodynamics and transport processes in Florida's Coral Reef}
The FRT is located along the northern side of the Straits of Florida, that connects the Gulf of Mexico (GoM) and the North Atlantic Ocean. The large-scale circulation in this region is dominated by the interplay of the Florida Current (FC) and the Loop Current (LC). The LC is an area of warm waters from the Caribbean Sea that enters the GoM through the Yucatan Channel and then loops anti-cyclonically through the northeastern Gulf of Mexico and finally exits through the southern Straits of Florida, where it forms the Florida Current (FC). The FC then flows northward between Florida and the Great Bahama Bank to form the Gulf Stream (Fig. \ref{intro:gom}). The northward penetration of the LC through the GoM varies throughout the year, and sporadically an anticyclonic ring separates from the current \citep{leipper1970sequence, maul1977annual, vukovich1988loop}. The detachment of these anticyclonic eddies from the main LC affects in turn the position of the FC, causing its meandering along the FRT. These variations of the FC position have been shown to have no seasonal pattern, and a strong interannual variability. The generated meandering is associated with the presence of cyclonic eddies between the core of the current and the FRT \citep{kourafalou2012florida}. These eddies significantly impact the ecology of the southwest Florida Shelf by causing larval retention on the shelf and delivering larval pulses to the Upper Keys \citep{lee1994evolution,limouzy1997translocation,sponaugle2005florida, kourafalou2012florida}. Furthermore, when the LC meets the shallow waters of the southwest corner of the WFS near the DRTO, it causes the inflow of nutrients that modify water properties on the shelf \citep{weisberg2003local, liu2016offshore}. These contacts have been hypothesized to impact in turn the penetration of the LC in the GoM \citep{weisberg2017loop}. However, on the northernmost section of the FRT, eddies generated by the meandering of the FC cause cold-water upwelling limiting the northern range of tropical species \citep{walker2013determining}. The circulation on the upper part of the shelf along the FRT is largely influenced by winds, as well as diurnal and semi-diurnal tidal processes \citep{lee2001transport, lee2002volume,d2007patterns}

>>>>

\section{Modeling hydrodynamics and transport processes in Florida's Coral Reef}
The FRT is located along the northern side of the Straits of Florida, that connects the Gulf of Mexico (GoM) and the North Atlantic Ocean. The large-scale circulation in this region is dominated by the interplay of the Florida Current (FC) and the Loop Current (LC). The LC is an area of warm waters from the Caribbean Sea that enters the GoM through the Yucatan Channel and then loops anti-cyclonically through the northeastern Gulf of Mexico and finally exits through the southern Straits of Florida, where it forms the Florida Current (FC). The FC then flows northward between Florida and the Great Bahama Bank to form the Gulf Stream (Fig. \ref{intro:gom}). The northward penetration of the LC through the GoM varies throughout the year, and sporadically an anticyclonic ring separates from the current \citep{leipper1970sequence, maul1977annual, vukovich1988loop}. The detachment of these anticyclonic eddies from the main LC affects in turn the position of the FC, causing its meandering along the FRT. These variations of the FC position have been shown to have no seasonal pattern, and a strong interannual variability. The generated meandering and \modif{its interaction with the topology of the Florida Keys} is associated with the presence of \modif{cold} cyclonic eddies between the core of the current and the FRT \citep{kourafalou2012florida}. These eddies significantly impact the ecology of the southwest Florida Shelf by causing larval retention on the shelf and delivering larval pulses to the Upper Keys \citep{lee1994evolution,limouzy1997translocation,sponaugle2005florida, kourafalou2012florida}. \modif{Furthermore, eddy-induced upwelling of deeper, nutrient-rich waters increase the chlorophyll and zooplankton concentrations in the Florida Keys \citep{lee1992influence}. Similarly, when the LC meets the shallow waters of the southwest corner of the WFS near the DRTO, it causes an inflow of nutrients supporting an ecosystem response \citep{weisberg2003local, liu2016offshore}}. These contacts have been hypothesized to impact in turn the penetration of the LC in the GoM \citep{weisberg2017loop}. However, on the northernmost section of the FRT, eddies generated by the meandering of the FC cause cold-water upwelling limiting the northern range of tropical species \citep{walker2013determining}. The circulation on the upper part of the shelf along the FRT is largely influenced by winds, as well as diurnal and semi-diurnal tidal processes \citep{lee2001transport, lee2002volume,d2007patterns}
